A non-additive generalization of probability theory \\for quantum mechanics and beyond

### 2025-08-08 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Turn equation into figure

\textbf{Fraction capacity.} Let us concentrate on the following problem: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$? How much of $\ens$ can be explained as coming from preparations corresponding to $A$? For example, let $\Ens$ be the space of all probability distributions for a six-sided die. Let $\ens_{123456}$ be the uniform distribution over all outcomes. Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability. Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$ (see \eqref{eq:dice} for an illustration), where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/5$ of $\ens$ can be constructed from $A$, but no more. Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more. Note that, given the uniform distribution, $1/5$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$. This gives the basic insight for our definitions.

\begin{equation}\label{eq:dice}
	\def\xstep{0.75em}
	\begin{tikzpicture}[y=3ex, x=\xstep]
		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/5) (2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
	\end{tikzpicture}
	= \frac{1}{6}
	\begin{tikzpicture}[y=3ex, x=\xstep]
		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1)};
	\end{tikzpicture}
	+ \frac{5}{6}
	\begin{tikzpicture}[y=3ex, x=\xstep]
		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
	\end{tikzpicture}
\end{equation}

>>>>

\textbf{Fraction capacity.} Let us concentrate on the following problem: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$? How much of $\ens$ can be explained as coming from preparations corresponding to $A$? For example, let $\Ens$ be the space of all probability distributions for a six-sided die. Let $\ens_{123456}$ be the uniform distribution over all outcomes. Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability. Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$ (see fig. \ref{fig:dice} for an illustration), where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/5$ of $\ens$ can be constructed from $A$, but no more. Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more. Note that, given the uniform distribution, $1/5$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$. This gives the basic insight for our definitions.

\begin{figure}
	\def\xstep{0.75em}
	\begin{tikzpicture}[y=3ex, x=\xstep]
		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/5) (2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
	\end{tikzpicture}
	$= \frac{1}{6}$
	\begin{tikzpicture}[y=3ex, x=\xstep]
		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1)};
	\end{tikzpicture}
	$+ \frac{5}{6}$
	\begin{tikzpicture}[y=3ex, x=\xstep]
		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
	\end{tikzpicture}
	\caption{Illustration of the mixture of distributions}\label{fig:dice}
\end{figure}
